---
title: "Survival Analysis"
author: "OA"
date: "December 1, 2015"
output: html_document
---
`library(survival)`
First create a survival object using `Surv()`
e.g.
`surv.obj<- Surv(addict$survt, addict$status == 1)`

Next, create a fit object using `survfit()` with surv.obj ~ 1 for no strata, or surv.obj ~ dat$strata 

plot e.g. N.B `conf.int=F` supresses 95% conf intervals
```{r eval=FALSE}
plot(surv.fit, lty = c("solid", "dashed"), col = c("black", "blue"), xlab = "survival time in days", ylab = "survival probabilities")
> legend("topright", c("Clinic 1", "Clinic 2"), lty = c("solid", "dashed"), col = c("black", "blue"))
```

Log-rank test e.g. this test looks at the observed versus expected events if the two curves were identical and compares them under a chi sq distribution
`survdiff(Surv(survt, status)~clinic, data = addict)`
OR
`survdiff(surv.obj ~ stratum)`

## Running a Cox PH model

`ph.model<-coxph(surv.obj ~ predictor1 + predictor2, data = data)`
then get a summary of the model object
`summary(ph.model)`

Next, check for interaction by creating a new model that contains interaction terms. The two models' log likelihood can be compared to see if there is significance to the full model. Use the script **liklihood ratio for two coxPH models.R**

## Checking the Cox PH assumption
### Graphical (Log-Log)
`plot(surv.fit, fun = "cloglog",xlab="time in days using logarithmic scale",ylab="log-log survival", main="log-log curves by category")`

### Schoenfeld residual correlation with survival time (significant p values indicate violation of the PH assumption)
`cox.zph(ph.model, transform = rank)`
`plot(cox.zph(ph.model, transform = rank), se = F, var = 'violated.variable')`

## Cox-adjusted survival curves
Need a survival object and a cox model based on the object. An adjusted survival curve is based on a specific pattern of data. e.g. for addict data set PRISON=0, DOSE=70, and CLINIC=2. Typically, the mean is picked for the covariates being adjusted for. All the covariates can be adjusted for by picking the mean for all covariates. *NOTE* this is better done with a stratified cox model and should be updated.

```{r eval = FALSE}
pattern1<- data.frame(clinic = 1, methodone = mean, prison = mean)
pattern2<- data.frame(clinic = 2, methodone = mean, prison = mean)
plot(surfit(ph.model, newdata=pattern1), conf.int = F)
par(new=T)
plot(surfit(ph.model, newdata=pattern2), conf.int = F)
par(new=F)
```

## Modeling Recurring Events
Data needs to be recorded in a counting process format with the following headers:

| id | interval | start | stop | event |
|--- | ---      | ---   | ---- | ---   |